Predicting the conformational preferences of proteins using a physics-based free energy method

### 2013-05-03 Paper/supporting_information.tex
changed 3May

We prepared computer generated models in both $\alpha$ and $\beta$ conformation for all five chamelon sequences mentioned 
in the main text. Crystallographic structure of GA95 and GB95 sequence have $\alpha$ and $\beta$ fold respectively. When,
we attempt to prepare the $\alpha$ and $\beta$ fold with GA95 sequence we start with the crystallographic 
structure of GA95 and GB95 with GA95 sequence. However, we remove the sidechains at three mutated positions in both the fold 
and call the program SCWRL4 \cite{Krivov2009} to generate the sidechain conformation of the mutated residues and its neighbors. Followed by 
this step, we do long molecular dynamics
minimization to remove any possible bad contact. This procedure was followed to generate all $\alpha$ and 
$\beta$ conformations of five sequences that are used for free energy calculation.

\begin{figure}
\includegraphics[width=5.6 in,height=5.3 in]{orban_perres_SI.pdf}
\label{fig:orban_full}
\caption{Per residue free energy calculation indicate significant differences of sidechain orientation of key amino acid residues. These amino acid
residues control the free energy equilibrium to either $\beta$ or
$\alpha$ conformation. All figures in the left are in $\beta$ conformation, whereas the right ones are in $\alpha$ conformation except in Figure (F).
Sidechains of hydrophobic residues Leu-7 in (A) and Ala-26 in (B) are oriented towards the protein interior in $\beta$ 
conformation. In such conformation they have hydrophobic interactions with other hydrophobic residues colored in white. In $\alpha$ conformation these residues are 
exposed to the protein surface. 
(C) Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $\beta$ conformation. (D) Gln-11 and Glu-15 form a H-bond in $\beta$ conformation which is absent
in the $\alpha$ conformation. (E) Hydrophobic Ile-49 is exposed to the protein surface in $\beta$ conformation, whereas it is inside protein hydrophobic core in
$\alpha$ conformation. (F) Sidechain of Ile-30 is inside the protein hydrophobic core in as it has a smaller sidechain with GA95 sequence in $\alpha$ conformation. 
In contrast, Phe-30 with larger sidechain is exposed to the protein surface with GB95 sequence.}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.6 in,height=3.4 in]{T0569.pdf}
\end{center}
\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569).
The best model was from the group "Mufold".}
\label{fig:T0569}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=5 in,height=4 in]{T0569_perres2.pdf}
\end{center}
\caption{Key differences between native crystal structure and the model as indicated by per residue free energy calculation.
The sidechain of hydrophobic residues, Val-59 and Ile-61 are oriented towards the protein hydrophobic core in (A) crystal 
structure, (B) exposed to the surfure in model. The beta sheet containing these residues is disordered in the generated model.
There is a H-bond between Lys-76 and Asp-11 in (D) model structure, which is absent in the (C) native structure.}
\label{fig:T0569_per_residue}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
\end{center}
\caption{Differences of the sidechain orientation in model 1 and crystallographic structure of CASP target T0538 as determined 
by per residue free energy calculation.  
(A) Two salt bridges between residues Glu-35-Arg-32 and Glu-28-Lys-24 exist in the native crystallographic structure. These are
absent in the model structure. Instead
(B) a new salt bridge is formed between residues Arg-32-Glu-28 in model. (C) Another salt bridge between
residues  Arg-26-Glu-50 in the native structure, (D) This salt bridge is absent in the model structure. In comparison to the crystallographic 
structure, orientation of Phe-51 differs in model1.}
\label{fig:T0538compare}
\end{figure}

\Ken{[This fig isn't very readable.  We need to focus into the sidechains, keeping only the nearby pieces of chain and leaving the rest out or making it fuzzy or something.  We need to see some point that you are making by looking at the figure.]}

@@

>>>>

We prepared computer generated models in both $\alpha$ and $\beta$ conformations for all five chamelon sequences mentioned 
in the main text. Crystallographic structure of GA95 and GB95 sequence have $\alpha$ and $\beta$ fold respectively. When
we attempt to prepare the $\alpha$ and $\beta$ fold with GA95 sequence, we start with the crystallographic 
structure of GA95 and GB95 with GA95 sequence. However, we remove the sidechains at three mutated positions in both the fold 
and call the program SCWRL4 \cite{Krivov2009} to generate the sidechain conformation of the mutated residues and its neighbors. Followed by 
this step, we do long molecular dynamics
minimization to remove any possible bad contact. This procedure was followed to generate all $\alpha$ and 
$\beta$ conformations of five sequences that are used for free energy calculation.

\setcounter{figure}{0}
\makeatletter 
\renewcommand{\thefigure}{S\@arabic\c@figure}
\begin{figure}
\includegraphics[width=5.6 in,height=5.3 in]{orban_perres_SI.pdf}
\label{fig:orban_full}
\caption{Per residue free energy calculation indicate significant differences of sidechain orientation of key amino acid residues. These amino acid
residues control the free energy equilibrium to either $\beta$ or
$\alpha$ conformation. All figures in the left are in $\beta$ conformation, whereas the right ones are in $\alpha$ conformation except in Figure (F) 
where both are in $\alpha$ conformation.
Sidechains of hydrophobic residues L7 in (A) and A26 in (B) are oriented towards the protein interior in $\beta$ 
conformation. In such conformation they have hydrophobic interactions with other hydrophobic residues colored in white. In $\alpha$ conformation these residues are 
exposed to the protein surface.  
(C) Y45 forms H-bond with D47 with GB95 sequence and $\beta$ conformation. (D) Q11 and E15 forms a salt bridge in $\beta$ conformation which is absent
in the $\alpha$ conformation. (E) Hydrophobic I49 is exposed to the protein surface in $\beta$ conformation, whereas in $\alpha$ conformation it 
is inside protein hydrophobic core. (F) Sidechain of I30 is inside the protein hydrophobic core as it has a smaller sidechain with GA95 sequence 
and $\alpha$ conformation. In contrast, Phe-30 with larger sidechain is exposed to the protein surface with GB95 sequence and $\beta$ conformation.}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5 in]{T0560.pdf}
    \end{center}
    \caption{Native structure and two models (from group ``Splicer'') for target T0560.}
    \label{fig:T0560}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.8 in,height=3.4 in]{T0569.pdf}
\end{center}
\caption{Native and best computer generated model structure of CASP Target T0569.} 
\label{fig:T0569}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=5 in,height=4 in]{T0569_perres2.pdf}
\end{center}
\caption{Key differences between experimental structure and the computer generated model model as indicated by per residue free energy calculation.
The sidechain of hydrophobic residues, V59 and I61 are oriented towards the protein hydrophobic core in (A)  
structure, (B) exposed to the surfure in model. The beta sheet containing these residues is disordered in the generated model.
A salt bridge between K76 and D11 in (D) computer generated structure, which is absent in the (C) native structure.}
\label{fig:T0569_per_residue}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{T0538.pdf}
\end{center}
\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP
    code:T0538). The model 1,2 and 3 are from the group PconsR, Shell and FOLDIT respectively.}
\label{fig:T0538}
\end{figure}

\section{What can we say about low resolution models from CASP experiments?}

In the main text, we discussed how well confinement method predicted preferences when the structures are not very far from the
native. But the question remains how far from native can we go and still see that the method produces correct result. In
this section we explored this question with models from  extracellular domain of the jumping translocation breakpoint
protein (pdb id: 2KJX). Most of the group could only generate low resolution models for this CASP9 target (id: T0531) .
In our comparison, we choose five models by the group MUFOLD-MD, which was the best performing group for this target
with their best model had a GDT\_TS value of 44. The result presented in Figure~\ref{fig:T0531} shows: 1.) native is
correctly identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and
the free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than 35  are ordered
incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free energies, meaning that
the actual ordering could change a lot \cite{Perez2012}. It is encouraging that at least the method can pick out the
best model even though it is got a low GDT score: 44.

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in]{T0531.pdf}
\end{center}
\caption{The native structure and 5 models of extracellular domain of the jumping translocation breakpoint protein (pdb
    id: 2KJX and the CASP code: T0531).}
\label{fig:T0531}
\end{figure}

@@

\bibitem{Perez2012}
Perez, A.; Yang, Z.; Bahar, I.; Dill, K.A.; MacCallum, J.L.; FlexE: Using Elastic Network Models to Compare Models of Protein Structure. J. Chem. Theory Comput., 2012, 8, 3985-3991.

### 2013-05-03 Paper/tt.tex
changed 3May

\Ken{[Guys---I recommend we get a better name than confinement.  It's totally uninformative.  In my view, better would
    be `confine-and-configure' or something.  Please give it some thought.]}

\Justin{[Ken---I agree that the name isn't great, but I don't think we should change it because: it's already in the
    literature and we didn't come up with the method. I think we should consider capitalizing it as Confinement Method
    throughout.]}

\section{Introduction}

@@

\Justin{[TI is actually the basis of the confinement step. We need to be more explicit about what we mean here.] Methods
    like thermodynamic integration~\cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free
    energy calculation may not be as useful for calculation of conformational free energy.} One widely explored strategy
is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. \Ken{[we need many refs here: cite
    Elber's milestoning, Jhih-Wei Chu, and the many other path-sampling approaches.]} The free energy along this
reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977}\Justin{[add ref
    where this is used computing free energy differences between conformations]} or targeted MD \Justin{[refs]}. Such
approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
conformations $A$ and $B$ are quite different, then it can be challenging to find such paths. Second, these methods are
computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires
accurate determinations of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$,
so each step requires substantial amounts of sampling.  And third, these methods are prone to large errors, because the
pathway error is a sum of many errors among the many steps.

\Ken{[Are all these methods below ones that don't involve paths?  Also, are there other key methods we should cite?  We
    need a good scholarly list of all the relevant works.]}

The calculation of protein conformational free energy has been successfully attempted by a number of groups
\cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
free energies that does not require knowing a pathway from $A$ to $B$.  \Ken{[please add the several new Karplus papers
    from 2012-2013.]}.

Here, we adopt the confinement method of Tyka et al~\cite{Tyka2006} and Cecchini et al.~\cite{Cecchini2009}, which is
based on the thermodynamic cycle shown in Figure~\ref{fig:method}. We start with an ensemble of related conformations,
$A$. We then restrict $A$ to a much ``tighter'' ensemble $A^\ast$ by applying position restraints in a series of MD
simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
$A^\ast$ and $B^\ast$ using either normal mode analysis or the quasiharmonic method [ref]. This step takes into account
both the remaining conformational entropy in each ensemble and the remaining enthalpy difference. In this way, we can
calculate the free energy difference between the two end states without needing to define a physical path or reaction
coordinate connecting the two states. The confinement method shares some similarities with the `confine-and-release'
method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006} \Ken{[cite mobley's jctc 07 (ref
    216 on our labsite), and mobley's recent j chem phys review, and mobley (ref 199) j chem phys, and others that are
    relevant.]}.

@@

\section{Results and Discussion}

\subsection{The confinement method succeeds at some basic consistency checks}

@@

Second, we looked at 6 target proteins from the CASP9 experiment [ref].  For each target, we examined up to 5 submitted
models.  We computed the difference free energy between the true native and the best model.
Figure~\ref{fig:summary_casp} and Table~\ref{table:casp_control} show that, in 5 out of 6 cases, the confinement method
assigns a lower free energy to the experimentally determined structure than to any of the decoys.  Other well-known
discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009} [add
rosetta holes, rosetta ff, dfire]; it is just an independent useful validation that the confinement calculations make
sense.

\Justin{[Arijit, please add GDT of best decoy to this table.]}

\begin{table}
\begin{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
    participants. We report the free energy difference between the most favorable decoy and the experimentally
    determined structure. Positive $\Delta\Delta G$ values indicate that the experimental structure is predicted to be
    more favorable than any of the decoys.}
\label{table:casp_control}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier & $\Delta \Delta G = G_{best~decoy} - G_{native}$ (kcal/mol) \\ \hline
     T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
     T0538       &    2L09        &          $-3.00 \pm 0.47$ \\ \hline
     T0540       &    3MX7        &          $16.94 \pm 0.49$ \\ \hline
     T0559       &    2L01        &          $2.10 \pm 0.24$ \\ \hline
     T0560       &    2L02        &          $18.00 \pm 0.49$ \\ \hline
     T0569       &    2KWY        &          $20.01 \pm 0.69$  \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{The confinement method correctly predicts the structures of chameleon sequences}

We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
al.!\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
sequence of 56-residues that is marginally stable in one of two possible folds. By mutating key residues in this
sequence they are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta +
\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the $\alpha$ conformation. One
pair of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair
(GA95/GB95) is 95 percent identical, differing in three positions. Accurately predicting the structural preferences of
these structures presents a serious challenge for computational
methods~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. \Justin{Do any of these citations try to
    do any modeling or comment about how challenging this would be?}

@@

\Ken{[ It's not clear what this test is and why it's better, and why there are only 5 sequences.  Please clarify.]}
\Justin{[There are five pairs of sequences that can be arranged into 3 different pairings. Arijit has removed the 5th
    sequence.]}

\Arijit{ Now there are 4 sequences as they have the experimentally available structure. The fifth sequence is discussed
    at the end of the paragraph.}

\Justin{[Ken: it's not that hard to tell an experimental structure from a computer generated model. Almost all of the
    knowledge-based scoring functions can do this a substantial fraction of the time. Comparing a near-native model with
    another model is a much more challenging problem.]}

@@

\subsection{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
\label{sec:orban_per_residue}

To better understand the mechanism behind these chameleon proteins, we decomposed the calculated free energy into
per-residue contributions in an approximate way~\cite{Tyka2006}. We decompose each confinement step ($\Delta
G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$) into per-residue contributions. We can also decompose the remaining enthalpy in
the confined ensemble into per-residue contributions. However, the method is approximate because we do not include the
residual conformational entropy from the normal mode or quasiharmonic analysis steps. Computing per-residue
contributions helps us to identify important residues that stabilize a particular conformation. The per residue free
energy, $\Delta \Delta G ((\beta) - (\alpha))$ is shown in Figure~\ref{fig:perresidue_orban}.

@@

\subsection{The confinement method is a useful tool for structure prediction}

@@

\subsubsection{The confinement method can correctly rank-order structure predictions}

@@

\Justin{[Arijit: Let's move the protein names, organisms, pdb ids, etc to a separate table and just use the CASP ids.]}

@@

\subsubsection{The per-residue free energy is sensitive to small changes in protein conformation}

In section~\protect\ref{sec:orban_per_residue} we discussed how the per-residue free energy can reveal mechanistic
detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
method and try to understand if the per-residue free energy can help us identify residues which stabilize or destabilize
a particular region of a protein. In this case, the two conformations have similar folds with only small changes in
localized regions. We chose CASP target T0569 and compared the experimental NMR structure with the best predicted model
(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the native structure is more stable
by 20 kcal/mol (Figure S2).

@@

\subsubsection{The confinement method occasionally produces incorrect results}

@@

\subsubsection{What can we say about low resolution models?}

@@

\item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
    $B^\ast$ using normal mode analysis or quasiharmonic analysis. The free energy calculated in this way is shown as
    $\Delta G_{A^\ast,B^\ast}$ in Figure~\ref{fig:method}.

@@

@@

\bibitem{Krivov2004}
Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, (41), 14766.

\bibitem{Alexander2007}Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and function. Proc. Natl. Acad. Sci. 2007, 104 (29), 11963-11968.

\bibitem{He2008}
He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105 (38), 14412-14417.

\bibitem{Alexander2009}
Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
, 106(50), 21149-21154.

\bibitem{Bryan2010}
Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20(4), 482-488.

\bibitem{He2012}
He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20(2), 2
83-91.

\bibitem{Shortle2009}
Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106(50), 21011-21012.

>>>>

\section*{Introduction}

@@

\Justin{[TI is actually the basis of the confinement step. We need to be more explicit about what we mean here.] Methods
    like thermodynamic integration~\cite{Tironi1994},\cite{Meirovitch2007} which are successful in alchemical free
    energy calculation may not be as useful for calculation of conformational free energy.} One widely explored strategy
is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006}. The free energy along this
reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977,Mascarenhas2013}  
and weighted histogram analysis method (WHAM) \cite{Kumar1992}. Such
approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
conformations $A$ and $B$ are quite different, then it can be challenging to find such paths. Second, these methods are
computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires
accurate determinations of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$,
so each step requires substantial amounts of sampling.  And third, these methods are prone to large errors, because the
pathway error is a sum of many errors among the many steps.

The calculation of protein conformational free energy has been successfully attempted by a number of groups
\cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
free energies that does not require knowing a pathway from $A$ to $B$.

Here, we adopt the confinement method of Tyka et al~\cite{Tyka2006} and Cecchini et al.~\cite{Cecchini2009}, which is
based on the thermodynamic cycle shown in Figure~\ref{fig:method}. We start with an ensemble of related conformations,
$A$. We then restrict $A$ to a much ``tighter'' ensemble $A^\ast$ by applying position restraints in a series of MD
simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
$A^\ast$ and $B^\ast$ using either normal mode analysis \cite{Brooks1983,Case1994} or the quasiharmonic 
method \cite{Karplus1981,Levy1984}. This step takes into account
both the remaining conformational entropy in each ensemble and the remaining enthalpy difference. In this way, we can
calculate the free energy difference between the two end states without needing to define a physical path or reaction
coordinate connecting the two states. The confinement method shares some similarities with the `confine-and-release'
method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006}.

@@

\section*{Results and Discussion}

\subsection*{The confinement method succeeds at some basic consistency checks}

@@

Second, we looked at 6 target proteins from the CASP9 experiment~\cite{Moult2011}.  For each target, we examined up to 5 submitted
models.  We computed the difference free energy between the true native and the best model.
Figure~\ref{fig:summary_casp} and Table~\ref{table:casp_control} show that, in 5 out of 6 cases, the confinement method
assigns a lower free energy to the experimentally determined structure than to any of the decoys.  Other well-known
discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009} [add
rosetta holes, rosetta ff, dfire]; it is just an independent useful validation that the confinement calculations make
sense.

\begin{table}
\begin{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
    participants. We report the free energy difference between the most favorable decoy and the experimentally
    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
    more favorable than any of the decoys.}
\label{table:casp_control}
\begin{tabular}{l l l l}\hline
    CASP Target  & PDB Identifier &  organism type  \\ \hline
     T0531       &    2KJX        &  extracellular domain of the jumping translocation breakpoint protein &  $44$ \\ \hline
     T0538       &    2L09        &  protein asr4154 from Nostoc sp. PCC7120        &  $96$ \\ \hline
     T0540       &    3MX7        &  human Fas apoptotic inhibitory molecule    &  $70$ \\ \hline
     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
     T0569       &    2KWY        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection*{The confinement method correctly predicts the structures of chameleon sequences}

We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012,Shortle2009}. These are instances in which two highly
similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
sequence of 56-residues that is marginally stable in one of two possible folds. By mutating key residues in this
sequence they are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta +
\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the $\alpha$ conformation. One
pair of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair
(GA95/GB95) is 95 percent identical, differing in three positions. Accurately predicting the structural preferences of
these structures presents a serious challenge for computational
methods \cite{Allison2011}.

@@

@@

\subsection*{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}
\label{sec:orban_per_residue}

To better understand the mechanism behind these chameleon proteins, we decomposed the calculated free energy into
per-residue contributions in an approximate way~\cite{Tyka2006}. We decompose each confinement step ($\Delta
G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$) into per-residue contributions. We can also decompose the remaining enthalpy in
the confined ensemble into per-residue contributions. However, the method is approximate because we do not include the
residual conformational entropy from the normal mode or quasiharmonic analysis steps. Computing per-residue
contributions helps us to identify important residues that stabilize a particular conformation. The per residue free
energy, $\Delta \Delta G (\beta - \alpha)$ is shown in Figure~\ref{fig:perresidue_orban}.

@@

\subsection*{The confinement method is a useful tool for structure prediction}

@@

\subsubsection*{The confinement method can correctly rank-order structure predictions}

@@

\begin{table}
\begin{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
    participants. We report the free energy difference between the most favorable decoy and the experimentally
    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
    more favorable than any of the decoys.}
\label{table:casp_control}
\begin{tabular}{l l l l}\hline
    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
     T0569       &    2KWY        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
\end{tabular}
\end{center}
\end{table}
\Justin{[Arijit: Let's move the protein names, organisms, pdb ids, etc to a separate table and just use the CASP ids.]}

@@

\subsubsection*{The per-residue free energy is sensitive to small changes in protein conformation:}

Previously, we discussed how the per-residue free energy can reveal mechanistic
detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
method and try to understand if the per-residue free energy can help us identify residues which stabilize or destabilize
a particular region of a protein. In this case, the two conformations have similar folds with only small changes in
localized regions. We chose CASP target T0569 and compared the experimental NMR structure with the best predicted model
(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the native structure is more stable
by 20 kcal/mol (Figure S2).

@@

\subsubsection*{The confinement method occasionally produces incorrect results:}

@@

\subsubsection*{What can we say about low resolution models?}

@@

\item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
    $B^\ast$ using normal mode analysis \cite{Brooks1983,Case1994} or quasiharmonic analysis \cite{Karplus1981,Levy1984}. The free energy calculated in this way is shown as
    $\Delta G_{A^\ast,B^\ast}$ in Figure~\ref{fig:method}.

@@

\bibitem{Mascarenhas2013}
Mascarenhas, N. M.; Kastner, J. How maltose influences structural changes to bind to maltose-binding protein: Results from umbrella sampling simulation. Proteins. 2013, 81, 185-198.

\bibitem{Kumar1992}
Kumar, S.; Rosenberg, J.M.; Bouzida, D.; Swendsen, R.H.; Kollman, P.A. The weighted histogram analysis method for free-energy calculations on biomolecules. I. The 
method. J Comput Chem. 1992, 13, 1011-1021.

@@

\bibitem{Krivov2004}
Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, 14766.

\bibitem{Brooks1983}
Brooks, B. R.; Karplus, M. Harmonic dynamics of proteins: Normal modes and fluctuations in bovine pancreatic trypsin inhibitor. Proc. Natl. Acad. Sci. U.S.A. 1983, 80, 6571-6575.

\bibitem{Case1994}
Case, D. Normal-mode analysis of protein dynamics Curr. Opin. Struct. Biol. 1994, 4, 285-290.

\bibitem{Karplus1981}
Karplus, M.; Kushick, J. Method for estimating the configurational entropy of macromolecules Macromolecules. 1981, 14, 325-332.

\bibitem{Levy1984}
Levy, R.; Karplus, M.; Kushick, J.; Perahia, D. Evaluation of the configurational entropy for proteins: application to molecular dynamics simulations of 
an $\alpha$-helix Macromolecules 1984, 17, 1370-1374.

\bibitem{Alexander2007}
Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and 
function. Proc. Natl. Acad. Sci. 2007, 104, 11963-11968.

\bibitem{He2008}
He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105, 14412-14417.

\bibitem{Alexander2009}
Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
, 106, 21149-21154.

\bibitem{Bryan2010}
Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20, 482-488.

\bibitem{He2012}
He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20,
83-91.

\bibitem{Shortle2009}
Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106, 21011-21012.

\bibitem{Allison2011}
Allison, J.R.; Bergeler, M.; van Gunsteren, W.F. Can computer modeling explain why two highly similar sequences fold into different structures ?
Biochemistry. 2011, 50, 10965-10973.
